A Fast and Tight Heuristic for A* in Road Networks

### 2021-03-12 paper/ch_potentials.tex
cleanup styles and use lipics 2021
address reviewer feedback

\documentclass[a4paper,USenglish,cleveref, autoref, thm-restate]{lipics-v2019}

@@

% \setcounter{secnumdepth}{2} %May be changed to 1 or 2 if section numbers are desired.

%This is a template for producing LIPIcs articles.
%See lipics-manual.pdf for further information.
%for A4 paper format use option "a4paper", for US-letter use option "letterpaper"
%for british hyphenation rules use option "UKenglish", for american hyphenation rules use option "USenglish"
%for section-numbered lemmas etc., use "numberwithinsect"
%for enabling cleveref support, use "cleveref"
%for enabling cleveref support, use "autoref"

@@

@@

%%
%% By default, the full list of authors will be used in the page
%% headers. Often, this list is too long, and will overlap
%% other information printed in the page headers. This command allows
%% the author to define a more concise list
%% of authors' names for this purpose.
% \renewcommand{\shortauthors}{Strasser and Zeitz}

%%
%% The abstract is a short summary of the work to be presented in the
%% article.
\begin{abstract}
We study exact, efficient and practical algorithms for route planning in large road networks.
Routing applications often require integrating the current traffic situation, planning ahead with traffic predictions for the future, respecting forbidden turns, and many other features depending on the exact application.
While Dijkstra's algorithm can be used to solve these problems, it is too slow for many applications.
A* is a classical approach to accelerate Dijkstra's algorithm.
A* can support many extended scenarios without much additional implementation complexity.
However, A*'s performance depends on the availability of a good heuristic that estimates distances.
Computing tight distance estimates is a challenge on its own.
On road networks, shortest paths can also be quickly computed using hierarchical speedup techniques.
They achieve speed and exactness but sacrifice A*'s flexibility.
Extending them to certain practical applications can be hard.
In this paper, we present an algorithm to efficiently extract distance estimates for A* from Contraction Hierarchies (CH), a hierarchical technique.
We call our heuristic CH-Potentials.
Our approach allows decoupling the supported extensions from the hierarchical speed-up technique.
Additionally, we describe A* optimizations to accelerate the processing of low degree nodes, which often occur in road networks.
\end{abstract}

@@

\section{Introduction}
\label{sec:intro}
The past decade has seen a plethora of research on route planning in large street networks~\cite{bdgmpsww-rptn-16}.
Routing a user through a road network can be formalized as the shortest path problem in weighted graphs.
Nodes represent intersections.
Roads are modeled using edges.
Edges are weighted by their traversal times.
The problem can be solved with Dijkstra's algorithm~\cite{d-ntpcg-59}.
Unfortunately, on continental sized networks, it is too slow for many applications.
Thus, speed-up techniques have been developed.
One popular example are Contraction Hierarchies~(CH)~\cite{gssv-erlrn-12}.
They have been used successfully in many real world applications.
A CH exploits the inherent hierarchy of road networks.
In a preprocessing step, additional shortcut edges are inserted, which allow skipping unimportant parts of the network at query time.
Another popular example is Multi-Level-Dijkstra~(MLD)~\cite{swz-umlgt-02} also known as CRP~\cite{dgpw-crprn-13}.
It is also used in practice~\cite{bingblog}.
MLD also uses shortcut edges.
% ich haette das at least eigentlich schon gerne drin. CH erreicht wesentlich mehr als 3 groessenordnungen
Both approaches achieve speed-ups of at least three orders of magnitude over Dijkstra's algorithm.

@@

Table~\ref{tab:building_blocks} depicts the performance of A* with different heuristics and optimizations.
We compare CH-Potentials to three other heuristics.
First, the Zero heuristic where $h(x)=0$ for all nodes $x$.
This corresponds to using Dijkstra's algorithm.
Second, we compare against our own implementation\footnote{
We use 16 landmarks generated with the avoid strategy~\cite{gw-cppsp-05} and activate all during every query.
Our ALT implementation is uni-directional.
In this work, we do not consider bidirectional search as it creates problems for some settings, such as predicted traffic.
} of ALT~\cite{gw-cppsp-05}.
Finally, we compare against a hypothetical \emph{Oracle-A*} heuristic.
This heuristic has instant access to a shortest distance array with respect to $w_\ell$, i.e. it is faster than the fastest heuristic possible in our model.
We fill this array before each query using a reverse Dijkstra search from the target node.
Thus, the reported running times of Oracle-A* do \emph{not} account for any heuristic evaluation.
Comparing against Oracle-A* allows us to measure the overhead of the heuristic evaluation of CH-Potentials.
Also, no other heuristic, which only has access to the preprocessing weights, can be faster than Oracle-A*.

@@

\bibliography{dblp,references}
\end{document}

>>>>

\documentclass[a4paper,UKenglish,cleveref, autoref, thm-restate]{lipics-v2021}

@@

@@

%Editor-only macros:: begin (do not touch as author)%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\EventEditors{David Coudert and Emanuele Natale}
\EventNoEds{2}
\EventLongTitle{19th International Symposium on Experimental Algorithms (SEA 2021)}
\EventShortTitle{SEA 2021}
\EventAcronym{SEA}
\EventYear{2021}
\EventDate{June 7--9, 2021}
\EventLocation{Nice, France}
\EventLogo{}
\SeriesVolume{190}
\ArticleNo{6}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\begin{abstract}
We study exact, efficient and practical algorithms for route planning in large road networks.
Routing applications often require integrating the current traffic situation, planning ahead with traffic predictions for the future, respecting forbidden turns, and many other features depending on the exact application.
While Dijkstra's algorithm can be used to solve these problems, it is too slow for many applications.
A* is a classical approach to accelerate Dijkstra's algorithm.
A* can support many extended scenarios without much additional implementation complexity.
However, A*'s performance depends on the availability of a good heuristic that estimates distances.
Computing tight distance estimates is a challenge on its own.
On road networks, shortest paths can also be quickly computed using hierarchical speedup techniques.
They achieve speed and exactness but sacrifice A*'s flexibility.
Extending them to certain practical applications can be hard.
In this paper, we present an algorithm to efficiently extract distance estimates for A* from Contraction Hierarchies (CH), a hierarchical technique.
We call our heuristic CH-Potentials.
Our approach allows decoupling the supported extensions from the hierarchical speed-up technique.
Additionally, we describe A* optimizations to accelerate the processing of low degree nodes, which often occur in road networks.
\end{abstract}

@@

\section{Introduction}
\label{sec:intro}
The past decade has seen a plethora of research on route planning in large street networks~\cite{bdgmpsww-rptn-16}.
Routing a user through a road network can be formalized as the shortest path problem in weighted graphs.
Nodes represent intersections.
Roads are modeled using edges.
Edges are weighted by their traversal times.
The problem can be solved with Dijkstra's algorithm~\cite{d-ntpcg-59}.
Unfortunately, on continental sized networks, it is too slow for many applications.
Thus, speed-up techniques have been developed.
One popular example are Contraction Hierarchies~(CH)~\cite{gssv-erlrn-12}.
They have been used successfully in many real world applications.
A CH exploits the inherent hierarchy of road networks.
In a preprocessing step, additional shortcut edges are inserted, which allow skipping unimportant parts of the network at query time.
Another popular example is Multi-Level-Dijkstra~(MLD)~\cite{swz-umlgt-02} also known as CRP~\cite{dgpw-crprn-13}.
It is also used in practice~\cite{bingblog}.
MLD also uses shortcut edges.
Both approaches achieve speed-ups of at least three orders of magnitude over Dijkstra's algorithm.

@@

Table~\ref{tab:building_blocks} depicts the performance of A* with different heuristics and optimizations.
We compare CH-Potentials to three other heuristics.
First, the Zero heuristic where $h(x)=0$ for all nodes $x$.
This corresponds to using Dijkstra's algorithm.
Second, we compare against our own implementation\footnote{
We use 16 landmarks generated with the avoid strategy~\cite{gw-cppsp-05} and activate all during every query.
Our ALT implementation is uni-directional.
In this work, we do not consider bidirectional search as it creates problems for some settings, such as predicted traffic.
} of ALT~\cite{gw-cppsp-05}.
Finally, we compare against a hypothetical \emph{Oracle-A*} heuristic.
This heuristic has instant access to a shortest distance array with respect to $w_\ell$, i.e. it is faster than the fastest heuristic possible in our model.
We fill this array before each query using a reverse Dijkstra search from the target node.
Thus, the reported running times of Oracle-A* do \emph{not} account for any heuristic evaluation.
CH-Potentials compute the same distance estimates but the heuristic evaluation has some overhead.
Comparing against Oracle-A* allows us to measure this overhead.
Also, no other heuristic, which only has access to the preprocessing weights, can be faster than Oracle-A*.

@@

\bibliography{references}
\end{document}
